---
title: "Run whole analysis"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```


## create folders
```{r}
dir.create("data")
dir.create("processing")
dir.create("results")
dir.create("results/figures")


dir.create("results/networks")
dir.create("results/networks/netDx")
dir.create("results/olink")
dir.create("results/clinical")
dir.create("results/LIMMA")
dir.create("results/figures/volcano_plots")
dir.create("results/GSEA_software")
dir.create("results/GSEA")
dir.create("results/Deseq2")
dir.create("results/ML")
dir.create("results/networks/BNN")
dir.create("results/networks/co_expression")
dir.create("results/GSEA_KEGG")
dir.create("results/MSEA")
dir.create("results/BINGO")
dir.create("results/figures/PCA")
dir.create("results/figures/boxplots_SASP")
dir.create("results/SCT")
dir.create("processing/MSEA")





dir.create("results/figures/boxplots")
dir.create("results/figures/boxplots/boxplots_Ale")


dir.create("results/GSMM")
dir.create("results/MATLAB")
dir.create("results/macrophages")
dir.create("results/figures/SCT")
dir.create("results/figures/SCT/gene_expression")
```

## run files in order
```{r}

```

## Preprocessing
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/clean/preprocessing/COCOMO_txn_met_preprocessing_2.Rmd
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/clean/preprocessing/2e_olink_analysis_preprocessing.Rmd
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/clean/preprocessing/extra_make_input_GSEA.Rmd

## clinical
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/clean/Clinical/2a_COCOMO_txn_Statistics.Rmd
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/clean/preprocessing/0_COCOMO_whole_cohort.Rmd



## SNF
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/SNF/1_COCOMO_txn_olink_metabolomics_SNF.Rmd
/home/flomik/Desktop/Code-PHD/COCOMO_txn/src/Whole_analysis/SNF/1_COCOMO_txn_olink_metabolomics_netDX.Rmd
