---
title: "Processing data macrophages"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

```{r}
library(xlsx)
library(mygene)
library(org.Hs.eg.db)
library(mygene)
library(ggplot2)
library(ggrepel)
library(limma)
library(ggpubr)
source("src/usefull_functions.R")
```

## proteomics
1) Normalization per batch
2) Batch correction
3) Filtering

Control vs Spermidine: Volcano plot marking some synaptic proteins and PSEA
HC-like plasma vs at-risk plasma adjusted for control : Volcano plot marking some synaptic proteins and PSEA

```{r}
col <- c("#793079",	"#999999",  "#797930", "#307930")
```

```{r}
cutoff_pval <- 0.05
```

### colors volcano
```{r}
cols <- c("up_regulated_>1.5" = "#cb4818",
          "up_regulated_1>x>1.5" = "#d27f28", 
          "up_regulated_0>x>1" = "#f1c232", 
          "down_regulated" = "#8fce00", 
          "NS" = "grey", 
          "significant" = "#586c6e",
          "down_regulated_>1.5" = "#274e13",
          "down_regulated_1>x>1.5" = "#6aa84f", 
          "down_regulated_0>x>1" = "#8fce00")
```

## load brain genes
```{r}
gns <- read.csv("data/brain_genes_clean_table.csv")
```


```{r}
name_ana <- "organoids_prot_frac_log2"
input <- "/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/Organoids/20230523_HF2_01_BrainOrganoids_TMT_fractions_proteins.csv"
```

```{r}
prot <- read.csv(input)
rownames(prot) <- prot$Accession
prot <- prot[colnames(prot) %in% colnames(prot)[grepl("Abundances\\.\\.", colnames(prot))]]
data <- data.frame(t(prot))
```

```{r}
table_samples <- data.frame(Name = rownames(data))
table_samples$extract <- gsub(".*F3\\.\\.", "", table_samples$Name)
table_samples$extract <- gsub("\\.\\..*", "", table_samples$extract)
table_samples$condition <- gsub(".*\\.\\.", "", table_samples$Name)
```



## remove empty lines
```{r}
print(ncol(data))
data <- data[,colSums(is.na(data)) < 10]
print(ncol(data))
```

### check for missing values
```{r}
counts <- 1:nrow(data) # create vector of appropriate length
for(i in 1:nrow(data)){
    # TRUE will be coerced to 1 for the summing
    counts[i] <- sum(is.na(data[i,]))
}
table(counts) # create the summary
```

### distribution
```{r}
data_2 <-as.matrix(data)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
```

### PCA before norm
```{r}
data$Name <- rownames(data)
pca_data <- merge(table_samples, data, by = "Name")
pca_data$extract <- NULL
```


```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-c(1:4)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```

```{r}
#df_out$sample <- as.factor(pca_data$sample)
```


```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)

ggsave(paste0("results/figures/PCA_before_norm_", name_ana, ".pdf"))
```
### log2 norm

```{r}
pca_data[,-c(1:2)] <- log2(1 + pca_data[,-c(1:2)])
data_norm <- pca_data[,-c(1:2)]
```

```{r}
path_results <- paste0("processing/", name_ana, "_norm_file.txt")
write.table(data_norm, file = path_results, 
            sep = "\t", row.names = TRUE, col.names = NA)
```
### Quantile normalization

```{r}
#x <- rownames(pca_data)
#y <- data.frame(t(pca_data[,-c(1:2)]))
#colnames(y) <- x
#path_data <- make_data_matrix(y, name_ana)
#path_design <- make_design_matrix(y, name_ana, pca_data$condition)
#outDir <- "processing/normalization"
```

```{r}
#data_norm <- read.delim("processing/normalization/organoids_prot_frac/organoids_prot_frac/median-normalized.txt", check.names = FALSE)
#rownames(data_norm) <- colnames(pca_data[,-c(1:2)])
```

### PCA after normalization
```{r}
pca_data <- data.frame(Name = pca_data$Name, condition = pca_data$condition, data_norm)
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-c(1:2)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```

```{r}
df_out$sample <- as.factor(pca_data$Name)
```

```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group, label = sample))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)+geom_text()

ggsave(paste0("results/figures/PCA_after_norm_", name_ana, ".pdf"))
```
```{r}
pca_data$condition[pca_data$Name == "Abundance..F3..127C..Sample..HC.like"] <- "At.risk"
pca_data$condition[pca_data$Name == "Abundance..F3..128N..Sample..At.risk"] <- "HC.like"
```
```{r}
pca_data$Name[pca_data$Name == "Abundance..F3..127C..Sample..HC.like"] <- "Abundance..F3..127C..Sample..At.risk_new"
pca_data$Name[pca_data$Name == "Abundance..F3..128N..Sample..At.risk"] <- "Abundance..F3..128N..Sample..HC.like_new"
```



```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-c(1:2)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```

```{r}
df_out$sample <- as.factor(pca_data$Name)
```

```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group, label = sample))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)+geom_text()

ggsave(paste0("results/figures/PCA_after_norm_", name_ana, "_change_sample.pdf"))
```

```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)
ggsave(paste0("results/figures/PCA_after_norm_", name_ana, "_change_sample_no_label.pdf"), width = 6, height = 5)
```


```{r}
pca_data_2 <- pca_data[pca_data$condition != "SPD",]
pca <- prcomp(pca_data_2[,-c(1:2)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data_2$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```

```{r}
df_out$sample <- as.factor(pca_data_2$Name)
```

```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)
ggsave(paste0("results/figures/PCA_after_norm_", name_ana, "_change_sample_no_label_without_SPD.pdf"), width = 6, height = 5)
```

### distribution after normalization
```{r}
data_2 <-as.matrix(data_norm)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
```

```{r}
write.csv(pca_data, paste0("processing/norm_data", name_ana, "_with_conditions_frac.csv"))
```

```{r}
pca_data <- pca_data[order(pca_data$condition),]
```


## Model 1 : Control vs Spermidine: Volcano plot marking some synaptic proteins and PSEA
```{r}
data <- pca_data[pca_data$condition %in% c("Control", "SPD"),]
```

```{r}
C <- factor(data$condition)

design <- model.matrix(~C)
design
```

```{r}
fit <- lmFit(t(data[,-c(1:2)]), design)
```

```{r}
fit.cont <- eBayes(fit)

top_table <- topTable(fit.cont, coef = 2, adjust="BH", n = Inf, sort.by = "P")
top_table$Comp <- "Control vs SPD"
top_table$BIOCHEMICAL <- rownames(top_table)
top_table <- top_table[,c(8, 1, 4, 5, 7)]


top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
top_table$synaptic_prot <- ifelse(top_table$Gene_name %in% gns$HumanName, "yes", NA)
path_table <- paste0("results/LIMMA/DGE_table_proteomics_organoids_SPD_controls_frac_log2.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < cutoff_pval, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_SPD_controls_frac_log2_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_SPD_controls_frac_log2_filt.xlsx")

#write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)

DF_top_table <- top_table[top_table$P.Value < 0.05, ]
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
print(nrow(DF_top_table[DF_top_table$logFC > 0,]))
print(nrow(DF_top_table[DF_top_table$logFC < 0,]))
```

## volcano plot
```{r}
dge <- top_table[,c(6, 2, 4, 3)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$P.Value < cutoff_pval, as.vector(tt$Gene_name), "")

#tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$P.Value > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"


vol <- ggplot(tt, aes(x = logFC, y = -log10(P.Value), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (P.Value)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_SPD_controls_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```

## Model 2: HC-like plasma vs at-risk plasma adjusted for control : Volcano plot marking some synaptic proteins and PSEA
```{r}
data <- pca_data[!pca_data$condition %in% c("SPD"),]
```

```{r}
C <- factor(data$condition, levels=c("Control","HC.like","At.risk"))
design <- model.matrix(~0+C)
colnames(design) <- levels(C)
fit <- lmFit(t(data[,-c(1:2)]), design)
```

```{r}
cont.matrix <- makeContrasts(Atriskvsctrl=At.risk-Control, 
                             Hclikevsctrl=HC.like-Control,
                             Hclikevsatrisk = At.risk-HC.like,
                              Diff=(At.risk-Control)-(HC.like-Control),
                                                      levels=design)
```

```{r}
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2)
```

```{r}

top_table <- topTable(fit2, coef = 1, adjust="BH", n = Inf, sort.by = "P")
top_table$Comp <- "Control vs at-risk"
top_table$BIOCHEMICAL <- rownames(top_table)
top_table <- top_table[,c(8, 1, 4, 5, 7)]

top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
top_table$synaptic_prot <- ifelse(top_table$Gene_name %in% gns$HumanName, "yes", NA)
path_table <- paste0("results/LIMMA/DGE_table_proteomics_organoids_at-risk_controls_frac_log2.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < cutoff_pval, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_at-risk_controls_frac_log2_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_at-risk_controls_frac_log2_filt.xlsx")

write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)

print(nrow(DF_top_table))
print(nrow(DF_top_table[DF_top_table$logFC > 0,]))
print(nrow(DF_top_table[DF_top_table$logFC < 0,]))
```
## volcano plot
```{r}
dge <- top_table[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$adj.P.Val < cutoff_pval, as.vector(tt$Gene_name), "")

#tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$adj.P.Val > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_at-risk_controls_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```


```{r}
my_comparisons <- list(c("HC.like", "At.risk"))
```

```{r}
data <- pca_data[pca_data$condition != "SPD",]
data$condition <- factor(data$condition, levels = c("Control", "HC.like", "At.risk"))
```

```{r}
a <- ggplot(data, aes(x = condition, y = P02776, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Condition", y = "PF4")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ scale_fill_manual(values=col[c(2, 3, 1)])+ 
  stat_compare_means(comparisons = my_comparisons)
  a
ggsave(paste0("results/figures/boxplot_PFA.pdf"), height = 3, width = 3.5)
```



```{r}
ggplot(pca_data, aes(y = Q9Y490,  x = condition)) + geom_point()
```

```{r}

top_table <- topTable(fit2, coef = 2, adjust="BH", n = Inf, sort.by = "P")
top_table$Comp <- "Control vs hc-like"
top_table$BIOCHEMICAL <- rownames(top_table)
top_table <- top_table[,c(8, 1, 4, 5, 7)]

top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
top_table$synaptic_prot <- ifelse(top_table$Gene_name %in% gns$HumanName, "yes", NA)
path_table <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_controls_frac_log2.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < cutoff_pval, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_controls_frac_log2_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_controls_frac_log2_filt.xlsx")

write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)

print(nrow(DF_top_table))
print(nrow(DF_top_table[DF_top_table$logFC > 0,]))
print(nrow(DF_top_table[DF_top_table$logFC < 0,]))
```
## volcano plot
```{r}
dge <- top_table[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$adj.P.Val < cutoff_pval, as.vector(tt$Gene_name), "")

#tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$adj.P.Val > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_hc-like_controls_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```

```{r}

top_table <- topTable(fit2, coef = 3, adjust="BH", n = Inf, sort.by = "P")
top_table$Comp <- "at-risk vs hc-like"
top_table$BIOCHEMICAL <- rownames(top_table)
top_table <- top_table[,c(8, 1, 4, 5, 7)]

top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
path_table <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_frac_log2.csv")
top_table$synaptic_prot <- ifelse(top_table$Gene_name %in% gns$HumanName, "yes", NA)
print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < cutoff_pval, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_frac_log2_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_frac_log2_filt.xlsx")

write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)

DF_top_table <- top_table[top_table$P.Value < 0.05, ]
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
print(nrow(DF_top_table[DF_top_table$logFC > 0,]))
print(nrow(DF_top_table[DF_top_table$logFC < 0,]))
```

## volcano plot
```{r}
dge <- top_table[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$adj.P.Val < cutoff_pval, as.vector(tt$Gene_name), "")

tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$adj.P.Val > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_hc-like_at-risk_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```


```{r}

top_table <- topTable(fit2, coef = 4, adjust="BH", n = Inf, sort.by = "P")
top_table$Comp <- "at-risk vs hc-like corrected"
top_table$BIOCHEMICAL <- rownames(top_table)
top_table <- top_table[,c(8, 1, 4, 5, 7)]
top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
top_table$synaptic_prot <- ifelse(top_table$Gene_name %in% gns$HumanName, "yes", NA)
path_table <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_corrected_frac_log2.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < cutoff_pval, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_corrected_frac_log2_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/DGE_table_proteomics_organoids_hc-like_at-risk_corrected_frac_log2_filt.xlsx")

write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)

DF_top_table <- top_table[top_table$P.Value < 0.05, ]
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
print(nrow(DF_top_table[DF_top_table$logFC > 0,]))
print(nrow(DF_top_table[DF_top_table$logFC < 0,]))
```

## volcano plot
```{r}
dge <- top_table[,c(6, 2, 4, 3)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA



for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$P.Value < cutoff_pval, as.vector(tt$Gene_name), "")

tt <- tt[tt$Gene_name != "PF4",]

#tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$P.Value > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"


vol <- ggplot(tt, aes(x = logFC, y = -log10(P.Value), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (P.Value)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_hc.like_at.risk_corrected_frac_log2_without_PF4.pdf")
ggsave(path_fig, width = 7, height = 4)
```

## volcano plot
```{r}
dge <- top_table[,c(6, 2, 4, 3)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA



for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$P.Value < cutoff_pval, as.vector(tt$Gene_name), "")

#tt <- tt[tt$Gene_name != "PF4",]

#tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$P.Value > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"


vol <- ggplot(tt, aes(x = logFC, y = -log10(P.Value), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (P.Value)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_hc.like_at.risk_corrected_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```

## volcano plot
```{r}
dge <- top_table[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < cutoff_pval){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > cutoff_pval){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 0 & tt$adj.P.Val < cutoff_pval, as.vector(tt$Gene_name), "")

tt$BIOCHEMICAL[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5] <- as.vector(tt$Gene_name)[tt$Gene_name %in% gns$HumanName & abs(tt$logFC) > 0.5]
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
tt$out[abs(tt$logFC) > 1.5 & tt$adj.P.Val > cutoff_pval] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(cutoff_pval), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_proteomics_organoids_hc-like_at-risk_corrected_frac_log2.pdf")
ggsave(path_fig, width = 7, height = 4)
```

