---
title: "Statistics clinical"
output: html_notebook
---


### clean environment
```{r}
rm(list=ls())
```


### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

```{r}
col <- c( "#8ea76b", "#6b8ea7",	"#a7846b")
```


```{r}
library(summarize)
library(dplyr)
library(xlsx)
library(ggplot2)
library(ggpubr)
library(reshape2)
```

```{r}
data_olink <- read.csv("processing/olink_data_new_ids.csv", row.names = 1, check.names = FALSE)
data_olink$Condition <- NULL
```

```{r}
clinical_2 <- read.csv("/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/clinical/data_base_200_patients_COCOMO_filtered.csv")
clinical_2 <- clinical_2[,colnames(clinical_2) %in% c("id", "diabetes")]
names(clinical_2)[1] <- "COCOMO_ID"
```

```{r}
clinical <- read.csv("processing/clinical_data_clean.csv", stringsAsFactors = FALSE)
clinical$X <- NULL
clinical <- clinical[!is.na(clinical$COCOMO_ID),]
clinical <- merge(clinical, clinical_2, by = "COCOMO_ID", all.x = TRUE)

cluster <- read.csv("processing/cluster_SNF_transcriptomics_metabolomics_olink_3.csv")
names(cluster)[3] <- "COCOMO_ID"
clinical <- merge(cluster, clinical, by = "COCOMO_ID", all.y = TRUE)
clinical$cluster[is.na(clinical$cluster) & clinical$HIV_Status == "Ctrl"] <- "Ctrl"

clinical$Mreads <- NULL
clinical$X..Q30 <- NULL
colnames(clinical) <- gsub("\\..*", "", colnames(clinical))
#clinical$User <- NULL
clinical$NGI <- NULL
```

```{r}
clinical <- clinical[clinical$COCOMO_ID %in% rownames(data_olink),]
```

```{r}
data <- clinical
data$AGE_cat[data$AGE < 55] <- "<55"
data$AGE_cat[data$AGE >= 55 & data$AGE < 65] <- "55<x<65"
data$AGE_cat[data$AGE >= 65] <- ">65"
data$AGE_cat <- factor(data$AGE_cat, levels = c("<55", "55<x<65", ">65"))
```

```{r}
table(data$Ethnic)
data$Ethnic[data$Ethnic != 1] <- 0

table(data$Origin)
data$Origin[data$Origin != 1] <- 0

table(data$Transm_mode)
data$Transm_mode[!data$Transm_mode %in% c(1,6) & !is.na(data$Transm_mode)] <- 0
```

```{r}
data$BMI_cat[data$BMI < 18.5] <- "underweight"
data$BMI_cat[data$BMI >= 18.5 & data$BMI < 25] <- "normal"
data$BMI_cat[data$BMI >= 25 & data$BMI < 30] <- "overweight"
data$BMI_cat[data$BMI >= 30 & data$BMI < 35] <- "obese"
data$BMI_cat[data$BMI >= 35] <- "extremely obese"

data$BMI_cat <- factor(data$BMI_cat, levels = c("underweight", "normal", "overweight", "obese", "extremely obese"))
```

```{r}
data$Combined_therapy <- gsub(".*ABC.*", "ABC", data$Combined_therapy)
data$Combined_therapy <- gsub(".*AZT.*", "AZT", data$Combined_therapy)
data$Combined_therapy <- gsub(".*TAF.*", "TAF-TDF", data$Combined_therapy)
data$Combined_therapy <- gsub(".*TDF.*", "TAF-TDF", data$Combined_therapy)
data$Combined_therapy[data$cluster != "Ctrl"] <- gsub(".*\\/.*", "Other", data$Combined_therapy[data$cluster != "Ctrl"])
data$Combined_therapy <- factor(data$Combined_therapy, levels = c("ABC", "AZT", "TAF-TDF", "Other"))
```

```{r}
table(data$Combined_therapy)
```

```{r}
x <- which(names(data) == "Combined_therapy")

names(data)[x] <- "first_drug"
data$first_drug
```

```{r}
data$Transm_mode[data$Transm_mode == 7] <- 3
data$Transm_mode[data$Transm_mode == 8] <- 3
```

```{r}
data$CDA_CD8 <- NA
data$CDA_CD8[!is.na(data$CD8)] <- data$CD4[!is.na(data$CD8)] / data$CD8[!is.na(data$CD8)]
```

```{r}
data$CDA_CD8_cat[!is.na(data$CDA_CD8)] <- ifelse(data$CDA_CD8[!is.na(data$CDA_CD8)] > 1, ">1", "<1")
data$CDA_CD8_cat <- as.factor(data$CDA_CD8_cat)
table(data$CDA_CD8_cat)
```


```{r}
data$Duration_cat <- NA
data$Duration_cat[data$Duration < 5] <- "<5"
data$Duration_cat[data$Duration >= 5 & data$Duration < 10] <- "5<x<10"
data$Duration_cat[data$Duration >= 10 & data$Duration < 15] <- "10<x<15"
data$Duration_cat[data$Duration >= 15] <- ">15"

data$Duration_cat <- as.factor(data$Duration_cat)
table(data$Duration_cat)
```

```{r}
data$VAT
data$VAT_cat <- NA
data$VAT_cat <- ifelse(data$VAT > 35 & data$GENDER == "Female", "high", data$VAT_cat)
data$VAT_cat <- ifelse(data$VAT < 35 & data$GENDER == "Female", "low", data$VAT_cat)
data$VAT_cat <- ifelse(data$VAT > 40 & data$GENDER == "Male", "high", data$VAT_cat)
data$VAT_cat <- ifelse(data$VAT < 40 & data$GENDER == "Male", "low", data$VAT_cat)
data$VAT_cat <- as.factor(data$VAT_cat)
table(data$VAT_cat)
```

## separate numeric and logical
```{r}
data$AGE <- as.numeric(data$AGE)
data$ob6 <- as.numeric(data$ob6)
data$ob10 <- as.numeric(data$ob10)

data$METS <- as.factor(data$METS)
data$central_obesity <- as.factor(data$central_obesity)
data$Ethnic <- as.factor(data$Ethnic)
data$Origin <- as.factor(data$Origin)
data$Transm_mode <- as.factor(data$Transm_mode)
data$hypertension <- as.factor(data$hypertension)
data$cluster <- as.factor(data$cluster)
data$diabetes <- NULL
data$X3rd_Drug <- as.factor(data$X3rd_Drug)
data$GENDER <- as.factor(data$GENDER)
```

```{r}
data$log10VL_ART <- log10(data$VL_ART)
```

```{r}
data$H2_54[data$H2_54 == 2] <- 1
data$H2_54[data$H2_54 == 7] <- 6
data$H2_54[data$H2_54 == 3] <- 4
data$H2_54 <- as.factor(data$H2_54)
```

```{r}
data$H2_55[data$H2_55 == 0] <- 1
data$H2_55[data$H2_55 == 4] <- 2
data$H2_55[data$H2_55 == 3] <- 2
data$H2_55[data$H2_55 == 6] <- 5
data$H2_55[data$H2_55 == 7] <- 5
data$H2_55 <- as.factor(data$H2_55)
```

```{r}
data$H2_56[data$H2_56 == 0] <- 1
data$H2_56[data$H2_56 == 4] <- 2
data$H2_56[data$H2_56 == 3] <- 2
data$H2_56[data$H2_56 == 6] <- 5
data$H2_56[data$H2_56 == 7] <- 5
data$H2_56 <- as.factor(data$H2_56)
```


```{r}
drug <- read.csv("/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/clinical/cocomo_drug_comorbidities_data.csv")
drug$X <- NULL
drug$ddiAZTd4t <- ifelse(drug$ddiAZTd4t == TRUE, 1 , 0)
colnames(drug)[1] <- "COCOMO_ID"
drug$alat <- as.numeric(drug$alat)
drug$creatini <- as.numeric(drug$creatini)
drug$ddiAZTd4t <- factor(drug$ddiAZTd4t)
```

```{r}
data <- merge(data, drug, by = "COCOMO_ID")
```

```{r}
# Function returning compact column classes
col_classes <- function(df) {
    t(as.data.frame(lapply(df, function(x) paste(class(x), collapse = ','))))
}

# Return example data's column classes

```

```{r}
data$X <- NULL
data$HIV_Status <- NULL
```

```{r}
clinical_ctrl <- data[,c(1:5, 38, 39)]
data_2 <- data
data <- data
write.csv(data, "processing/clinical_data_clean_with_clusters_and_categories.csv")
data <- data[data$cluster != "Ctrl",]
```

```{r}
data$diabetes <- as.factor(data$diabetes)
```

```{r}
data$X <- NULL
data$User <- NULL
data_factor <- data[, c(which(col_classes(data) == "factor"))]
data_numeric <- data[, c(2, which(col_classes(data) != "factor"))]
```

```{r}
names(data_factor)
```
```{r}
names(data_numeric)
```

### Numeric test normality

```{r}
library(nortest)
#library(MVN)
```
data_numeric
## test normality
```{r}
j = 1
list_norm <- list()
for (i in 3:ncol(data_numeric)){
  c1 <- data_numeric[data_numeric$cluster == 1,i]
  c2 <- data_numeric[data_numeric$cluster == 2,i]
  #c3 <- data_numeric[data_numeric$cluster == 3,i]
  #c4 <- data_numeric[data_numeric$cluster == 4,i]
  #c5 <- data_numeric[data_numeric$cluster == 5,i]
  test_c1 <- as.numeric(shapiro.test(c1)$p.value)
  test_c2 <- as.numeric(shapiro.test(c2)$p.value)
  #test_c3 <- as.numeric(shapiro.test(c3)$p.value)
  #test_c4 <- as.numeric(shapiro.test(c4)$p.value)
  #test_c5 <- as.numeric(shapiro.test(c5)$p.value)
  test_c3 <- 0.01
  test_c4 <- 0.01
  test_c5 <- 0.01
  #print(c(names(data_numeric)[i], test_c1, test_c2, test_c3, test_c4, test_c5))
  if(test_c1 < 0.05 & test_c2 < 0.05 & test_c3 < 0.05 & test_c4 < 0.05 & test_c5 < 0.05){
    print(paste0(names(data_numeric)[i], "is following normal distribution"))
    list_norm[[j]] <- names(data_numeric)[i]
    j = j + 1
  }
}
```
## normally distributed
```{r}
data_norm <- data_numeric[,colnames(data_numeric) %in% c( "cluster", unlist(list_norm))]
```

## global test
```{r}
table_numeric_2 <- data.frame(parameter = rep(names(data_norm)[2:ncol(data_norm)], each = 2), cluster = rep(c(1, 2), length(names(data_norm)[2:ncol(data_norm)])), Mean_SD = NA, pval_anova = NA)
j = 1

for (i in 2:ncol(data_norm)){
  c1 <- data_norm[data_norm$cluster == 1,i]
  c2 <- data_norm[data_norm$cluster == 2,i]
  #c3 <- data_norm[data_norm$cluster == 3,i]
  #c4 <- data_norm[data_norm$cluster == 4,i]
  #c5 <- data_norm[data_norm$cluster == 5,i]
  x <- paste0("Mean =",round(mean(c1, na.rm=TRUE), 2), ", SD =", round(sd(c1, na.rm = TRUE), 2))
  y <- paste0("Mean =",round(mean(c2, na.rm=TRUE), 2), ", SD =", round(sd(c2, na.rm = TRUE), 2))
  #z <- paste0("Mean =",round(mean(c3, na.rm=TRUE), 2), ", SD =", round(sd(c3, na.rm = TRUE), 2))
  #a <- paste0("Mean =",round(mean(c4, na.rm=TRUE), 2), ", SD =", round(sd(c4, na.rm = TRUE), 2))
  #b <- paste0("Mean =",round(mean(c5, na.rm=TRUE), 2), ", SD =", round(sd(c5, na.rm = TRUE), 2))
  table_numeric_2[c(j, j+1), 3] <- c(x, y)
  test <- t.test(c1, c2)
  table_numeric_2[j, 4] <- test$p.value
  j = j + 2
}


table_numeric_2$X <- paste0("C", table_numeric_2$cluster, "_", table_numeric_2$parameter)

table_numeric_2$X <- NULL

write.xlsx(table_numeric_2, "results/clinical/COCOMO_3_layers_normal_distrib_numeric_values_ANOVA_3.xlsx")
```

```{r}
data_numeric$COCOMO_ID <- NULL
```

```{r}
medIQR <- function(x, na.rm = TRUE, ...) {
  qq <- stats::quantile(x, c(0.50, 0.25, 0.75), na.rm = na.rm, ...)
  structure(qq, 
            names = c("Median", "Q1", "Q3"),
            class = "medIQR") # how about EstInterval? smry?
}

medIQR.formula <- function(formula,
                           data = NULL, ...,
                           subset, na.action = NULL,
                           overall = FALSE,
                           overall.label = "Overall") {
  if (missing(formula) || (length(formula) != 3)) 
    stop("'formula' missing or incorrect")
  m <- match.call(expand.dots = FALSE)
  if (is.matrix(eval(m$data, parent.frame()))) 
    m$data <- as.data.frame(data)
  m$... <- m$overall <- m$overall.label <- NULL
  m$na.action <- na.action
  require(stats, quietly = TRUE)
  m[[1L]] <- as.name("model.frame")
  mf <- eval(m, parent.frame())
  response <- attr(attr(mf, "terms"), "response")
  s <- split(mf[[response]], mf[-response])
  
  if(overall) {
    if(any(names(s) == overall.label))
    stop(sQuote(overall.label), " is an existing level")
    s <- c(Overall = list(unlist(s)), s)
  }
  structure(sapply(s, FUN = medIQR, ...),
            class = "medIQR")
}
```

```{r}
table_numeric <- data.frame(parameter = rep(names(data_numeric)[2:ncol(data_numeric)], each = 2), cluster = rep(c(1, 2), length(names(data_numeric)[2:ncol(data_numeric)])), Mean_SD = NA, pval_anova = NA)
j = 1

for (i in 2:ncol(data_numeric)){
  c1 <- data_numeric[data_numeric$cluster == 1,i]
  c2 <- data_numeric[data_numeric$cluster == 2,i]
  #c3 <- data_numeric[data_numeric$cluster == 3,i]
  #c4 <- data_numeric[data_numeric$cluster == 4,i]
  #c5 <- data_numeric[data_numeric$cluster == 5,i]
  x <- paste0(round(medIQR(c1)[[1]], 2), "(", round(medIQR(c1)[[2]], 2), "-", round(medIQR(c1)[[3]], 2), ")")
  y <- paste0(round(medIQR(c2)[[1]], 2), "(", round(medIQR(c2)[[2]], 2), "-", round(medIQR(c2)[[3]], 2), ")")
  #z <- paste0("Mean =",round(mean(c3, na.rm=TRUE), 2), ", SD =", round(sd(c3, na.rm = TRUE), 2))
  #a <- paste0("Mean =",round(mean(c4, na.rm=TRUE), 2), ", SD =", round(sd(c4, na.rm = TRUE), 2))
  #b <- paste0("Mean =",round(mean(c5, na.rm=TRUE), 2), ", SD =", round(sd(c5, na.rm = TRUE), 2))
  table_numeric[c(j, j+1), 3] <- c(x, y)
  test <- wilcox.test(c1, c2)
  print(test)
  table_numeric[j, 4] <- test$p.value
  j = j + 2
}




table_numeric$X <- paste0("C", table_numeric$cluster, "_", table_numeric$parameter)

#table_numeric <-  merge(table_numeric, reg, by = "X")

table_numeric$X <- NULL

write.xlsx(table_numeric, "results/clinical/COCOMO_3_layers_non_norm_numeric_values_kruskal_3.xlsx")
```

```{r}
table(data$cluster)
```

```{r}
data_factor$cluster <- as.factor(as.vector(data_factor$cluster))
```

```{r}
table_factor <- data.frame(Var1 = NA, Var2 = NA, Level = NA, Freq= NA, Freq2 = NA, Parameter = NA, test = NA, pval = NA)
j = 1

for (i in 2:ncol(data_factor)) {
  print(names(data_factor)[i])
  x <- data.frame(t(table(data_factor$cluster, data_factor[,i])))
  x$Level <- x$Var1
  x$Var1 <- paste0(names(data_factor)[i], "_", x$Var1, "_", x$Var2)
  x$Freq2 <- ""
  x$Freq2[x$Var2 == 1] <- paste0(x$Freq[x$Var2 == 1], "(", round(x$Freq[x$Var2 == 1]/sum(x$Freq[x$Var2 == 1])*100, 2), "%)")
  x$Freq2[x$Var2 == 2] <- paste0(x$Freq[x$Var2 == 2], "(", round(x$Freq[x$Var2 == 2]/sum(x$Freq[x$Var2 == 2])*100, 2), "%)")
  #x$Freq2[x$Var2 == 3] <- paste0(x$Freq[x$Var2 == 3], "(", round(x$Freq[x$Var2 == 3]/sum(x$Freq[x$Var2 == 3])*100, 2), "%)")
  #x$Freq2[x$Var2 == 4] <- paste0(x$Freq[x$Var2 == 4], "(", round(x$Freq[x$Var2 == 4]/sum(x$Freq[x$Var2 == 4])*100, 2), "%)")
  #x$Freq2[x$Var2 == 5] <- paste0(x$Freq[x$Var2 == 5], "(", round(x$Freq[x$Var2 == 5]/sum(x$Freq[x$Var2 == 5])*100, 2), "%)")
  x$Parameter <- names(data_factor)[i]
  m1 <- rbind(table(data_factor$cluster, data_factor[,i]))
  f <- chisq.test(m1)
  print(f$expected)
  if(sum(as.vector(f$expected > 5), na.rm = TRUE) == length(as.vector(f$expected))){
    x[1, 7] <- "Chi-Square Test"
    print(x[1, 7])
    x[1, 8] <- f$p.value
    }else{
      x[1, 7] <- "Fisher's Exact Test"
      print(x[1, 7])
      x[1, 8] <- fisher.test(m1, simulate.p.value=TRUE)$p.value
    }
  names(x)[7:8] <- c("test", "pval")
  table_factor <- rbind(x,table_factor)
}

table_factor <- table_factor[,colnames(table_factor) %in% c("Var2", "Parameter", "Level", "Freq2", "test", "pval")]
names(table_factor)[1] <- "cluster"
write.xlsx(table_factor, "results/clinical/X2_results_results_cocomo_3_groups_3.xlsx")
```
```{r}
x <- table_numeric$parameter[table_numeric$pval_anova < 0.05 & !is.na(table_numeric$pval_anova)]
x
```

```{r}
table_numeric_2$parameter[table_numeric_2$pval_anova < 0.05 & !is.na(table_numeric_2$pval_anova)]
```


```{r}
y <- table_factor$Parameter[table_factor$pval < 0.05 & !is.na(table_factor$pval)]
y
```

```{r}
write.csv(y, "processing/factor_differing_groups_3.csv")
write.csv(x, "processing/numeric_differing_groups_3.csv")
```

## Duration
```{r}
data$cluster <- factor(data$cluster, levels = c("Ctrl", 1, 2))
num_data <- data[, colnames(data) %in% c("cluster", x)]
log_data <- data[, colnames(data) %in% c("cluster", y)]
```

## boxplot numeric data

for (r in 2:ncol(num_data)) {
  name <- colnames(num_data)[r]
  x <- pairwise.wilcox.test(num_data[,r], num_data$cluster, p.adj = "none")
  x2 <- data.frame(x$p.value)
  colnames(x2) <- gsub("X", "", colnames(x2))
  colnames(x2)
  list_pairs <- list()
  a <- 1
  for (i in 1:nrow(x2)) {
    for (j in 1:ncol(x2)) {
      if(!is.na(x2[i, j]) & x2[i, j] < 0.05){
        list_pairs[[a]] <- c(colnames(x2)[j], rownames(x2)[i])
        a = a + 1
      }
    }
  }
  a <- ggplot(num_data, aes(x = as.factor(cluster), y = num_data[,r], fill = as.factor(cluster), color =as.factor(cluster))) + 
    geom_boxplot(position=position_dodge(1), size = 0.5, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Cluster", y = colnames(num_data)[r])+ 
    theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x = element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="bottom",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + scale_fill_manual(values=col[2:3])+ scale_color_manual(values=col[2:3]) + 
    geom_jitter(shape=16, size=2, color="black", position=position_jitter(0.05))+ stat_compare_means(comparisons = list_pairs, method = "wilcox.test", label = "p.signif", bracket.size = 0.8, size = 5)
  ggsave(paste0("results/figures/boxplot_", colnames(num_data)[r], "_3.pdf"))
  
}


## logical
```{r}
pairwise.chisq.test <- function(x, g, p.adjust.method = p.adjust.methods, ...) {
  DNAME <- paste(deparse(substitute(x)), "and", deparse(substitute(g)))
  g <- factor(g)
  p.adjust.method <- match.arg(p.adjust.method)

  compare.levels <- function(i, j) {
    xi <- x[as.integer(g) == i]
    xj <- x[as.integer(g) == j]
    m1 <- table(xi)
    m2 <- table(xj)
    m <- rbind(m1, m2)
    print(m)
    chisq.test(m)$p.value
  }
  PVAL <- pairwise.table(compare.levels, levels(g), p.adjust.method)
  ans <- list(method = "chi-squared test", data.name = DNAME, p.value = PVAL, 
              p.adjust.method = p.adjust.method)
  class(ans) <- "pairwise.htest"
  ans
}
```

## barplot logical data

log_data <- log_data[log_data$cluster != "Ctrl",]
log_data$cluster <- as.factor(as.vector(log_data$cluster))

for (r in 2:ncol(log_data)) {
  name <- colnames(log_data)[r]
  cluster <- as.vector(log_data$cluster)
  comp <- as.vector(log_data[,r])
  x <- pairwise.chisq.test(comp, cluster, p.adjust.method = "none")
  
  x2 <- data.frame(x$p.value)
  colnames(x2) <- gsub("X", "", colnames(x2))
  colnames(x2)
  list_pairs <- list()
  a <- 1
  for (i in 1:nrow(x2)) {
    for (j in 1:ncol(x2)) {
      if(!is.na(x2[i, j]) & x2[i, j] < 0.05){
        list_pairs[[a]] <- c(colnames(x2)[j], rownames(x2)[i])
        a = a + 1
      }
    }
  }
  print(list_pairs)
  m1 <- data.frame(rbind(table(cluster, comp)))
  m1$cluster <- rownames(m1)
  m1 <- melt(m1)
  p <- ggplot(m1, aes(x=cluster, y=value, fill=variable)) + 
    geom_bar(stat="identity")+theme_minimal()
  p
  ggsave(paste0("results/figures/barplot_", colnames(log_data)[r], "_3.pdf"))
  
}

## PCAs
(labelled for BMI x, duration of treatment (0à< 5,  5 à< 10, 10 à15, +15)x, genderx, CD4/CD8 ratio (< 1 and > 1)x, METSx, VATx,  transmission modex, 3rd drug x)
```{r}
#clinical_pca <- select(data_2, COCOMO_ID, cluster, BMI_cat, Duration_cat, GENDER, CDA_CD8_cat, METS, VAT_cat, Transm_mode, X3rd_Drug)
```

```{r}
#write.csv(clinical_pca, "processing/clinical_data_PCA_3.csv")
```

