---
title: "Processing data macrophages"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

```{r}
library(xlsx)
library(mygene)
library(org.Hs.eg.db)
library(mygene)
library(ggplot2)
library(ggrepel)
library(limma)
source("src/usefull_functions.R")
```

## proteomics
1) Normalization per batch
2) Batch correction
3) Filtering

```{r}
name_ana <- "spermine_prot_norm_frac"
input <- "/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/proteomics/Ujjwal_4Donors_TMT16_Frac.csv"
```

```{r}
prot <- read.csv(input)
rownames(prot) <- prot$Accession
prot <- prot[colnames(prot) %in% colnames(prot)[grepl("Abundances\\.\\.", colnames(prot))]]
data <- data.frame(t(prot))
```

```{r}
table_samples <- data.frame(Name = rownames(data))
table_samples$extract <- gsub(".*F21\\.\\.", "", table_samples$Name)
table_samples$extract <- gsub("\\.\\..*", "", table_samples$extract)
```

```{r}
table_samples$donor <- NA
table_samples$donor[table_samples$extract %in% c("126", "129N", "129C", "131C")] <- "D1"
table_samples$donor[table_samples$extract %in% c("130C", "127N", "127C", "132C")] <- "D2"
table_samples$donor[table_samples$extract %in% c("130N", "132N", "128N", "128C")] <- "D3"
table_samples$donor[table_samples$extract %in% c("131N", "133N", "133C", "134N")] <- "D4"

table_samples$condition <- gsub(".*F21\\.\\.", "", table_samples$Name)
table_samples$condition <- gsub(".*\\.\\.", "", table_samples$condition)
table_samples$samples <- paste0(table_samples$condition, "_", table_samples$donor)
```

```{r}
cell <- read.xlsx("/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/proteomics/Copy of CellNumbers_Sabrina.xlsx", 1)
cell <- cell[,c(4, 6)]
names(cell)[1:2] <- c("Cell_nb","extract")

table_samples <- merge(table_samples, cell, by = "extract")
```

## remove empty lines
```{r}
print(ncol(data))
data <- data[,colSums(is.na(data)) < 10]
print(ncol(data))
```

### check for missing values
```{r}
counts <- 1:nrow(data) # create vector of appropriate length
for(i in 1:nrow(data)){
    # TRUE will be coerced to 1 for the summing
    counts[i] <- sum(is.na(data[i,]))
}
table(counts) # create the summary
```

### distribution
```{r}
data_2 <-as.matrix(data)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
```
```{r}
table_condition <- table_samples
table_condition$extract <- NULL
```

### PCA before norm
```{r}
data$Name <- rownames(data)
pca_data <- merge(table_condition, data, by = "Name")
rownames(pca_data) <- pca_data$Name_2
pca_data$Name <- NULL
```

```{r}
emptycols <- colSums(is.na(pca_data)) == nrow(pca_data)
pca_data <- pca_data[!emptycols]
```

```{r}
pca_data[is.na(pca_data)] <-  0
#pca_data[,-c(1:4)] <- pca_data[,-c(1:4)]/pca_data[, 4]
  

pca <- prcomp(pca_data[,-c(1:4)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```
+ scale_color_manual(values = col)+ scale_fill_manual(values = col)

```{r}
df_out$sample <- as.factor(pca_data$condition)
```
```{r}
col <- c( "#005900", "#878782",	"#761212")
```

```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group, label = sample))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+geom_text(hjust=0.5, vjust=2)

ggsave(paste0("results/figures/PCA_before_norm_", name_ana, ".pdf"))
```

### Quantile normalization

```{r}
x <- rownames(pca_data)
y <- data.frame(t(pca_data[,-c(1:4)]))
colnames(y) <- x
path_data <- make_data_matrix(y, name_ana)
path_design <- make_design_matrix(y, name_ana, pca_data$samples)
outDir <- "processing/normalization"
```

```{r}
data_norm <- read.delim("processing/normalization/Prot_spermine/Prot_spermine/median-normalized.txt", check.names = FALSE)
rownames(data_norm) <- colnames(pca_data[,-c(1:4)])
```

```{r}
path_results <- paste0("processing/", name_ana, "_norm_file.txt")
write.table(data_norm, file = path_results, 
            sep = "\t", row.names = TRUE, col.names = NA)
```

```{r}
pca_data <- data.frame(condition = pca_data$condition, sample = pca_data$samples, replicate = pca_data$donor, cell = pca_data$Cell_nb, t(data_norm))
```

### PCA after normalization
```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-c(1:4)])
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$condition)
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
```

```{r}
df_out$sample <- as.factor(pca_data$condition)
```

+ scale_color_manual(values = col)+ scale_fill_manual(values = col)
```{r}
ggplot(df_out, aes( x= PC1, y = PC2, color = group, label = sample))+ theme_classic()+geom_point(size = 5, alpha = 0.9, shape = 19)+ 
        theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
              legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
              axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
              axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        xlab(paste0("PC1: ",round(a1,0),"% variance")) +
        ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
          theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
                axis.title.y = element_text(size=16, face = "bold",color="black"),
                axis.text = element_text(color="black",size=16, face = "bold"),
                axis.text.x =element_text(color="black",size=16, face = "bold"),
                panel.border = element_blank(),
                axis.ticks.x = element_blank(),
                strip.text.x = element_text(size = 9,color="black"),
                panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
                plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
                legend.background=element_rect(fill = alpha("white", 0)))+geom_text(hjust=0.5, vjust=2)

ggsave(paste0("results/figures/PCA_after_norm_", name_ana, ".pdf"))
```

### distribution after normalization
```{r}
data_2 <-as.matrix(data_norm)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
```

```{r}
write.csv(pca_data, "processing/norm_data_with_conditions_norm_data_log2_frac.csv")
```

> lev <- c("wt.0hr","wt.6hr","wt.24hr","mu.0hr","mu.6hr","mu.24hr")
> f <- factor(targets$Target, levels=lev)
> design <- model.matrix(~0+f)
> colnames(design) <- lev
> fit <- lmFit(eset, design)

```{r}
pca_data <- pca_data[order(pca_data$replicate),]
pca_data <- pca_data[order(pca_data$condition),]
C <- factor(pca_data$condition, levels = c("Control", "Spermine", "Spermidine", "Both"))
R <- factor(pca_data$replicate)
Cell <- pca_data$cell
design <- model.matrix(~ 0 + Cell + C + R)
design
colnames(design)[1]
```

```{r}
fit <- lmFit(t(pca_data[,-c(1:4)]), design)
```

```{r}
cont.matrix = makeContrasts(
  spermidine_control = CSpermidine-CControl,
  spermine_control = CSpermine-CControl,
  both_control = CBoth-CControl,
  levels=design)

vfit <- contrasts.fit(fit, contrasts=cont.matrix)
fit.cont <- eBayes(vfit)
```

```{r}
comp <- c("CSpermidine-CControl", "CSpermine-CControl", "CBoth-CControl")
```

```{r}
top_table <- data.frame(BIOCHEMICAL = NA, logFC= NA, P.Value= NA, adj.P.Val= NA, Comp= NA)

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- top_table_1[,c(8, 1, 4, 5, 7)]
  top_table <- rbind(top_table, top_table_1)
  top_table_1$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table_1$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
  write.csv(top_table_1[, c(6, 2, 3)], paste0("processing/GSEA/proteomics_spermin_frac_", comp[i],".csv"), quote = FALSE, row.names = F)
}

top_table <- top_table[complete.cases(top_table),]
top_table$Gene_name <- mapIds(org.Hs.eg.db, as.vector(top_table$BIOCHEMICAL), keytype="UNIPROT", column="SYMBOL")
path_table <- paste0("results/LIMMA/LIMMA_spermin_proteomics_frac.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.1, ]
print(nrow(DF_top_table))
DF_top_table <- DF_top_table[!is.na(DF_top_table$Gene_name),]
print(nrow(DF_top_table))
#DF_top_table <- DF_top_table[!duplicated(DF_top_table$Gene_name),]
#print(nrow(DF_top_table))

path_results <- paste0("results/LIMMA/LIMMA_spermin_proteomics_frac_filt.csv")

write.table(DF_top_table, file = path_results)

path_results <- paste0("results/LIMMA/LIMMA_spermin_proteomics_frac_filt.xlsx")

write.xlsx(DF_top_table, file = path_results)

table(DF_top_table$Comp)
```
```{r}
ggplot(pca_data, aes(y = O60242,  x = condition)) + geom_point()
```
```{r}
top_table_1 <- top_table[top_table$Comp == "CSpermidine-CControl",]
top_table_2 <- top_table[top_table$Comp == "CSpermine-CControl",]
top_table_3 <- top_table[top_table$Comp == "CBoth-CControl",]
```

## ven diagram R
```{r}
a <- top_table_1$Gene_name[!is.na(top_table_1$Gene_name) & top_table_1$adj.P.Val < 0.1]
b <- top_table_2$Gene_name[!is.na(top_table_2$Gene_name) & top_table_2$adj.P.Val < 0.1]
c <- top_table_3$Gene_name[!is.na(top_table_3$Gene_name) & top_table_3$adj.P.Val < 0.1]
```

```{r}
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")
```

```{r}
library(VennDiagram)
# Chart
venn.diagram(
  x = list(a, b, c),
  category.names = c("Spermidine vs Ctl" , "Spermine vs Ctl" , "Spermine + Spermidine vs Ctrl"),
  filename = 'results/figures/spermidine_spermine_proteomics_quantile.pdf',
  output=TRUE,
  fill = myCol,
  cat.default.pos = "text"
)
```
```{r}
mac <- read.xlsx("data/macrophages/M1 and M2 marker gene list_AG Savai_March 2022.xlsx", 1)
mac_M1 <- as.vector(mac$Gene.Symbol)
mac_M1 <- gsub(" \\(.*", "", mac_M1)

mac_M1[mac_M1 == "ACOD1/IRG1"] <- "ACOD1"
mac_M1[mac_M1 == "NOS2\\iNOS"] <- "NOS2"
mac_M1[mac_M1 == "CXCL8\\IL8"] <- "CXCL8"
mac_M1 <- c("IL8", "IRG1", mac_M1)

mac_M2 <- as.vector(mac$Gene.Symbol.1)
mac_M2[mac_M2 == "MRC1\\CD206"] <- "MRC1"
mac_M2 <- c("CD206", mac_M2)

M <- c(mac_M1, mac_M2)
```


## volcano plot
```{r}
dge <- top_table_1[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA


for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > 0.1){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% c("CCL5", "IL1B", "CD163", "SCARB2", "CXCL8", "HLA-DRB1", "STAT1") & !is.na(tt$Gene_name), as.vector(tt$Gene_name), NA)
#tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) > 1 & tt$adj.P.Val < 0.1 | tt$adj.P.Val < 0.007 , as.vector(tt$Gene_name), NA)
#tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), as.vector(tt$Gene_name), tt$BIOCHEMICAL)
tt$M <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), "M", NA)
tt$out[abs(tt$logFC) > 2.5 & tt$adj.P.Val > 0.1] <- "out"           
tt <- tt[is.na(tt$out), ]

length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])


tt$color[is.na(tt$color)] <- "significant"
#tt$Gene_name[tt$color == "significant"] <- NA
cols <- c("up_regulated_>1.5" = "#cb4818",
          "up_regulated_1>x>1.5" = "#d27f28", 
          "up_regulated_0>x>1" = "#f1c232", 
          "down_regulated" = "#86b7e8", 
          "NS" = "grey", 
          "significant" = "#586c6e",
          "down_regulated_>1.5" = "#35495c",
          "down_regulated_1>x>1.5" = "#5d80a2", 
          "down_regulated_0>x>1" = "#86b7e8")

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(0.1), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_norm_CSpermidine-CControl_frac_quantile_6_biomarkers.pdf")
ggsave(path_fig, width = 10, height = 6)
```
## volcano plot
```{r}
dge <- top_table_2[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > 0.1){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% c("CCL5", "IL1B", "CD163", "SCARB2", "CXCL8", "HLA-DRB1", "STAT1", "STAT3") & !is.na(tt$Gene_name), as.vector(tt$Gene_name), NA)

#tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) >1 & tt$adj.P.Val < 0.1 | tt$adj.P.Val < 0.007 , as.vector(tt$Gene_name), NA)
                         
length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])
#tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), as.vector(tt$Gene_name), tt$BIOCHEMICAL)
tt$M <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), "M", NA)
tt$out[abs(tt$logFC) > 2.5 & tt$adj.P.Val > 0.1] <- "out"           
tt <- tt[is.na(tt$out), ]


tt$color[is.na(tt$color)] <- "significant"
#tt$Gene_name[tt$color == "significant"] <- NA
cols <- c("up_regulated_>1.5" = "#cb4818",
          "up_regulated_1>x>1.5" = "#d27f28", 
          "up_regulated_0>x>1" = "#f1c232", 
          "down_regulated" = "#86b7e8", 
          "NS" = "grey", 
          "significant" = "#586c6e",
          "down_regulated_>1.5" = "#35495c",
          "down_regulated_1>x>1.5" = "#5d80a2", 
          "down_regulated_0>x>1" = "#86b7e8")

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_norm_CSpermine-CControl_frac_quantile_6_biomarkers.pdf")
ggsave(path_fig, width = 10, height = 6)
```
## volcano plot
```{r}
dge <- top_table_3[,c(6, 2, 3, 4)]
tt <- dge

tt$X <- NULL
name_ana <- unique(tt$Comp)
tt$color <- NA

for (i in 1:nrow(tt)){
  if(tt[i, 2] > 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_>1.5"
  }
  if(tt[i, 2] > 1 & tt[i, 2] < 1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_1>x>1.5"
  }
  if(tt[i, 2] > 0 & tt[i, 2] < 1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "up_regulated_0>x>1"
  }
  if(tt[i, 2] < -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_>1.5"
  }
  if(tt[i, 2] < -1 & tt[i, 2] > -1.5 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_1>x>1.5"
  }
  if(tt[i, 2] < 0 & tt[i, 2] > -1 & tt[i, 4] < 0.1){
    tt[i, ncol(tt)] <- "down_regulated_0>x>1"
  }
  if(tt[i, 4] > 0.1){
    tt[i, ncol(tt)] <- "NS"
    #tt[i, 1] <- NA
  }
}

tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% c("CCL5", "IL1B", "CD163", "SCARB2", "CXCL8", "HLA-DRB1", "STAT1") & !is.na(tt$Gene_name), as.vector(tt$Gene_name), NA)
#tt$BIOCHEMICAL <- ifelse(abs(tt$logFC) >1 & tt$adj.P.Val < 0.1 | tt$adj.P.Val < 0.007 , as.vector(tt$Gene_name), NA)

length(tt$BIOCHEMICAL[!is.na(tt$BIOCHEMICAL)])

#tt$BIOCHEMICAL <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), as.vector(tt$Gene_name), tt$BIOCHEMICAL)
tt$M <- ifelse(tt$Gene_name %in% M & !is.na(tt$Gene_name), "M", NA)
tt$out[abs(tt$logFC) > 2 & tt$adj.P.Val > 0.1] <- "out"           
tt <- tt[is.na(tt$out), ]

tt$color[is.na(tt$color)] <- "significant"
#tt$Gene_name[tt$color == "significant"] <- NA
cols <- c("up_regulated_>1.5" = "#cb4818",
          "up_regulated_1>x>1.5" = "#d27f28", 
          "up_regulated_0>x>1" = "#f1c232", 
          "down_regulated" = "#86b7e8", 
          "NS" = "grey", 
          "significant" = "#586c6e",
          "down_regulated_>1.5" = "#35495c",
          "down_regulated_1>x>1.5" = "#5d80a2", 
          "down_regulated_0>x>1" = "#86b7e8")

vol <- ggplot(tt, aes(x = logFC, y = -log10(adj.P.Val), colour = color, size = abs(logFC)))+ 
  geom_point(alpha = 1, na.rm = T)+ 
  theme(legend.title=element_text(size=2),legend.text=element_text(size=2),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=25),axis.title.x=element_text(size=25),
        axis.text.y=element_text(size=25),axis.text.x=element_text(size=25)) +
  labs(x="Log2 Fold Change",y="-log10 (adj.P.Val)")+
  geom_text_repel(aes(label = BIOCHEMICAL),size=2,label.size = 0.1, color = "black", xlim = c(NA, Inf),
    # Do not repel from top or bottom edges.
    ylim = c(-Inf, Inf))+ theme_classic()+ 
  geom_hline(yintercept=-log10(0.1), linetype="dashed", color = "black")+scale_color_manual(values = cols)

print(vol)
path_fig <- paste0("results/figures/volcano_plot_norm_CBoth-CControl_frac_quantile__biomarkers.pdf")
ggsave(path_fig, width = 10, height = 6)
```

```{r}
mac <- read.xlsx("data/macrophages/M1 and M2 marker gene list_AG Savai_March 2022.xlsx", 1)
mac_M1 <- as.vector(mac$Gene.Symbol)
mac_M1 <- gsub(" \\(.*", "", mac_M1)

mac_M1[mac_M1 == "ACOD1/IRG1"] <- "ACOD1"
mac_M1[mac_M1 == "NOS2\\iNOS"] <- "NOS2"
mac_M1[mac_M1 == "CXCL8\\IL8"] <- "CXCL8"
mac_M1 <- c("IL8", "IRG1", mac_M1)

mac_M2 <- as.vector(mac$Gene.Symbol.1)
mac_M2[mac_M2 == "MRC1\\CD206"] <- "MRC1"
mac_M2 <- c("CD206", mac_M2)
```

```{r}
top_table_1 <- top_table[top_table$Comp == "CSpermidine-CControl",]
top_table_1 <- top_table_1[!duplicated(top_table_1$Gene_name),]
top_table_2 <- top_table[top_table$Comp == "CSpermine-CControl",]
top_table_2 <- top_table_2[!duplicated(top_table_2$Gene_name),]
top_table_3 <- top_table[top_table$Comp == "CBoth-CControl",]
top_table_3 <- top_table_3[!duplicated(top_table_3$Gene_name),]
```

## heatmap M1 and M2 markers
```{r}
rownames(pca_data) <- pca_data$sample
data <- data.frame(t(pca_data[,-c(1:4)]))
data$BIOCHEMICAL <- rownames(data)

top_table_1 <- top_table_1[,c(1, 2, 4, 6)]
names(top_table_1)[2:3] <- c("LFC_Spermidine_Ctrl", "FDR_Spermidine_Ctrl")
top_table_2 <- top_table_2[,c(1, 2, 4)]
names(top_table_2)[2:3] <- c("LFC_Spermine_Ctrl", "FDR_Spermine_Ctrl")
top_table_3 <- top_table_3[,c(1, 2, 4)]
names(top_table_3)[2:3] <- c("LFC_Both_Ctrl", "FDR_Both_Ctrl")

data <- merge(top_table_1, data, by = "BIOCHEMICAL")
data <- merge(top_table_2, data, by = "BIOCHEMICAL")
data <- merge(top_table_3, data, by = "BIOCHEMICAL")

data$macrophage <- ifelse(data$Gene_name %in% mac_M1, "M1", NA)
data$macrophage <- ifelse(data$Gene_name %in% mac_M2, "M2", data$macrophage)
data <- data[!is.na(data$macrophage),]

data$LFC_Both_Ctrl[data$FDR_Both_Ctrl > 0.1] <- NA
data$LFC_Spermine_Ctrl[data$FDR_Spermine_Ctrl > 0.1] <- NA
data$LFC_Spermidine_Ctrl[data$FDR_Spermidine_Ctrl > 0.1] <- NA
```

```{r}
data$D1_both <- data$Both_D1-data$Control_D1
data$D1_spermine <- data$Spermine_D1-data$Control_D1
data$D1_spermidine <- data$Spermidine_D1-data$Control_D1

data$D2_both <- data$Both_D2-data$Control_D2
data$D2_spermine <- data$Spermine_D2-data$Control_D2
data$D2_spermidine <- data$Spermidine_D2-data$Control_D2

data$D3_both <- data$Both_D3-data$Control_D3
data$D3_spermine <- data$Spermine_D3-data$Control_D3
data$D3_spermidine <- data$Spermidine_D3-data$Control_D3

data$D4_both <- data$Both_D4-data$Control_D4
data$D4_spermine <- data$Spermine_D4-data$Control_D4
data$D4_spermidine <- data$Spermidine_D4-data$Control_D4
```

```{r}
library(circlize)
library(ComplexHeatmap)
col_fun1 = colorRamp2(c(-1.5,-1, -0.5, 0, 0.5, 1, 1.5), c("#1e90ff","#2676d0" ,"#275ea2","black","#a39a21","#d0c521","#fff21e"))
col_fun_2 = colorRamp2(c(-1, 0, 1), c("blue", "white", "red"))
```

```{r}
rownames(data) <- data$Gene_name
```

```{r}
library(viridisLite)
```

```{r}
col = viridis(100)
```

```{r}
condition <- colnames(data)[-c(1:25)]
condition <- gsub("D.\\_", "", condition)
condition <- factor(condition, levels = rev(c("both", "spermidine", "spermine")))
```

col_fun1 = colorRamp2(c(-1,-0.1, -0.01, 0, 0.01,0.1, 1), c("#1e90ff","#2676d0" ,"#275ea2","black","#a39a21","#d0c521","#fff21e"))

```{r}
write.csv(data, "processing/data_foldchange_spermine_spermidine.csv")
```

```{r}
data <- data[!is.na(data$LFC_Spermidine_Ctrl),]
```

```{r}
data <- data[order(data$macrophage, data$LFC_Spermidine_Ctrl),]
```

```{r}
pdf("results/figures/Heatmap_proteomics_spermine_spermidine_markers_macrophages.pdf", width = 7, height = 4)
ht = Heatmap(data[,-c(1:25)], 
             column_order = sort(colnames(data[,-c(1:25)])),
             row_order = rownames(data),
             col = col_fun1, 
             show_column_names = TRUE, 
             show_row_names = TRUE,
             column_split = condition, 
             #row_split = data_2$under_group, 
             #column_title = c("Ctrl",1, 2), 
             #column_title_gp = gpar(fill = col), 
             border = FALSE, 
             column_gap = unit(2, "mm"), 
             right_annotation = rowAnnotation(
                                              LogFC_spermine = data$LFC_Spermine_Ctrl,
                                              LogFC_spermidine = data$LFC_Spermidine_Ctrl,
                                              LogFC_both = data$LFC_Both_Ctrl,

                                              Marker = data$macrophage,
                                              col = list(LogFC_both = col_fun_2,
                                                         LogFC_spermidine = col_fun_2,
                                                         LogFC_spermine = col_fun_2), show_legend = TRUE))
draw(ht)
dev.off()

ht
```



### genes qPCR spermine / spermidine
```{r}
sperm <- read.xlsx("/home/flomik/Desktop/Code-PHD/COCOMO_txn/manuscript/spermidine_genes.xlsx", 1)
sperm$name_2 <- as.vector(sperm$gene)
sperm$uniprot <- mapIds(org.Hs.eg.db, as.vector(sperm$gene), keytype="SYMBOL", column="UNIPROT")
sperm$name_2[sperm$gene == "VPAT"] <- "SLC18B1"
sperm$name_2[sperm$gene == "MAT1"] <- "MNAT1"
sperm$name_2[sperm$gene == "DNMT"] <- "DNMT1"
sperm$name_2[sperm$gene == "Mtase"] <- "DNMT3B"
sperm$uniprot <- mapIds(org.Hs.eg.db, as.vector(sperm$name_2), keytype="SYMBOL", column="UNIPROT")
```

```{r}
x <- top_table$BIOCHEMICAL[top_table$Gene_name %in% sperm$name_2 & !is.na(top_table$Gene_name)]
```

```{r}
sperm_2 <- top_table[top_table$Gene_name %in% sperm$name_2 & !is.na(top_table$Gene_name) | top_table$BIOCHEMICAL %in% sperm$uniprot & !is.na(top_table$Gene_name) ,]
```

```{r}
write.csv(sperm_2, "manuscript/proteins_spermine_pathways_detected.csv")
write.csv(sperm, "manuscript/proteins_spermine_list.csv")
```


```{r}
pca_data <- pca_data[,colnames(pca_data) %in% c("sample", "condition", "replicate", sperm$uniprot, x)]
```

```{r}
colnames(pca_data)
```
```{r}
sperm_2$Gene_name[sperm_2$BIOCHEMICAL == "P53396"]
sperm_2$Gene_name[sperm_2$BIOCHEMICAL == "P52788"]
sperm_2$Gene_name[sperm_2$BIOCHEMICAL == "P31153"]
sperm_2$Gene_name[sperm_2$BIOCHEMICAL == "P19623"]
sperm_2$Gene_name[sperm_2$BIOCHEMICAL == "P05089"]
```

```{r}
pca_data$condition <- factor(pca_data$condition, levels = c("Control", "Spermine", "Spermidine", "Both"))
```

## boxplot
```{r}
a <- ggplot(pca_data, aes(x = condition, y = P31153, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Group", y = "MAT2A")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  a
  ggsave(paste0("results/figures/boxplots/MAT2A_boxplot_proteins_spermidine_spermine.pdf"), height = 4, width = 3)
```

## boxplot
```{r}
a <- ggplot(pca_data, aes(x = condition, y = P19623, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Group", y = "SRM")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  a
  ggsave(paste0("results/figures/boxplots/SRM_boxplot_proteins_spermidine_spermine.pdf"), height = 4, width = 3)
```
## boxplot
```{r}
a <- ggplot(pca_data, aes(x = condition, y = P05089, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Group", y = "ARG1")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  a
  ggsave(paste0("results/figures/boxplots/ARG1_boxplot_proteins_spermidine_spermine.pdf"), height = 4, width = 3)
```

## boxplot
```{r}
a <- ggplot(pca_data, aes(x = condition, y = P53396, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Group", y = "ACLY")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  a
  ggsave(paste0("results/figures/boxplots/ACLY_boxplot_proteins_spermidine_spermine.pdf"), height = 4, width = 3)
```

## boxplot
```{r}
a <- ggplot(pca_data, aes(x = condition, y = P52788, fill = condition)) + 
    geom_boxplot(position=position_dodge(1), size = 0.1, alpha =0.8, outlier.shape = NA)+ 
    labs(x="Group", y = "SMS")+ 
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          legend.position="none",
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0))) + 
    geom_jitter(shape=16, size=1, color="black", position=position_jitter(0.05))+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  a
  ggsave(paste0("results/figures/boxplots/SMS_boxplot_proteins_spermidine_spermine.pdf"), height = 4, width = 3)
```

