---
title: "SNF transcriptomics / metabolomics / olink"
output: html_notebook
---


## Aim
Clustering patients based on omics profile

### clean environment
```{r}
rm(list=ls())
```

### set directory (replace by your directory)
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

### load packages
```{r}
library(DESeq2)
library(dplyr)
library(SNFtool)
library(xlsx)
library(ggplot2)
library(corrr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(circlize)
```


```{r}
col_c <- c("#6b8ea7",	"#a7846b")
```

## load data
### transcriptomics
```{r}
data_txn <- read.csv("processing/COCOMO_transcriptomics_vsd_data_deseq2.csv", row.names = 1, check.names = FALSE)
#data_txn$COCOMO_ID <- rownames(data_txn)
```
### metabolomics
```{r}
data_met <- read.csv("processing/COCOMO_metabolomics_log2_norm_filt.csv", row.names = 1, check.names = FALSE)
data_met <- data_met[rownames(data_met) %in% rownames(data_txn), ]
data_met <- data_met[order(match(rownames(data_met), rownames(data_txn))),]
#data_met$COCOMO_ID <- rownames(data_met)
```

### olink
```{r}
data_olink <- read.csv("processing/olink_data_new_ids.csv", row.names = 1, check.names = FALSE)
data_olink$Condition <- NULL
```


### clinical parameters
```{r}
clinical <- read.csv("processing/clinical_data_clean_2.csv")
clinical$X <- NULL
#clinical <- clinical[,c(1, 4, 6, 8, 11, 12, 13, 36, 37, 38, 40)]
clinical_HIV <- clinical[clinical$HIV_Status != "Ctrl",]
#clinical_HIV <- clinical
clinical_HIV <- clinical_HIV[clinical_HIV$COCOMO_ID %in% rownames(data_olink),]
```

```{r}
table(clinical$HIV_Status)
```

```{r}
table(clinical_HIV$HIV_Status)
```

```{r}
data_txn <- data_txn[rownames(data_txn) %in% clinical_HIV$COCOMO_ID,]
data_met <- data_met[rownames(data_met) %in% clinical_HIV$COCOMO_ID,]
data_olink <- data_olink[rownames(data_olink) %in% clinical_HIV$COCOMO_ID,]
```

```{r}
data_met <- data_met[order(match(rownames(data_met), rownames(data_txn))),]
data_olink <- data_olink[order(match(rownames(data_olink), rownames(data_txn))),]
```

## filter data
```{r}
data_met <- t(data_met)
data_met <- data_met[apply(data_met, 1, var) >= 0.2, ]
data_met <- t(data_met)

data_olink <- t(data_olink)
data_olink <- data_olink[apply(data_olink, 1, var) >= 0.2, ]
data_olink <- t(data_olink)

data_txn <- t(data_txn)
data_txn <- data_txn[apply(data_txn, 1, var) >= 0.2, ]
data_txn <- t(data_txn)
```

## Normalization
```{r}
Data1 = standardNormalization(data_txn)
Data2 = standardNormalization(data_met)
Data3 = standardNormalization(data_olink)
```

## Distance matrices
```{r}
Dist1 = dist2(as.matrix(Data1),as.matrix(Data1));
Dist2 = dist2(as.matrix(Data2),as.matrix(Data2));
Dist3 = dist2(as.matrix(Data3),as.matrix(Data3));
```

```{r}
.discretisation <- function(eigenVectors) {
  
  normalize <- function(x) x / sqrt(sum(x^2))
  eigenVectors = t(apply(eigenVectors,1,normalize))
  
  n = nrow(eigenVectors)
  k = ncol(eigenVectors)
  
  R = matrix(0,k,k)
  R[,1] = t(eigenVectors[round(n/2),])
  
  mini <- function(x) {
    i = which(x == min(x))
    return(i[1])
  }
  
  c = matrix(0,n,1)
  for (j in 2:k) {
    c = c + abs(eigenVectors %*% matrix(R[,j-1],k,1))
    i = mini(c)
    R[,j] = t(eigenVectors[i,])
  }
  
  lastObjectiveValue = 0
  for (i in 1:20) {
    eigenDiscrete = .discretisationEigenVectorData(eigenVectors %*% R)
    
    svde = svd(t(eigenDiscrete) %*% eigenVectors)
    U = svde[['u']]
    V = svde[['v']]
    S = svde[['d']]
    
    NcutValue = 2 * (n-sum(S))
    if(abs(NcutValue - lastObjectiveValue) < .Machine$double.eps) 
      break
    
    lastObjectiveValue = NcutValue
    R = V %*% t(U)
    
  }
  
  return(list(discrete=eigenDiscrete,continuous =eigenVectors))
}  
```

```{r}
.discretisationEigenVectorData <- function(eigenVector) {
  
  Y = matrix(0,nrow(eigenVector),ncol(eigenVector))
  maxi <- function(x) {
    i = which(x == max(x))
    return(i[1])
  }
  j = apply(eigenVector,1,maxi)
  Y[cbind(1:nrow(eigenVector),j)] = 1
  
  return(Y)
  
}
```

```{r}
my_estimateNumberOfClustersGivenGraph <- function(W, NUMC=2:5) {
    # Estimates the best number of clusters from a vector of choices, using 
    #   the eigen-gap & rotation cost heuristics.
    #
    # Args:
    #   W: Affinity matrix (usually result from SNF)
    #   NUMC: A vector of integers specifying which cluster numbers to check
    #
    # Returns:
    #   A vector of the top two suggested number of clusters using
    #       the eigen-gap and rotation cost heuristics. 
    #

    #Put this check after the length(NUMC) check?
    if (min(NUMC) == 1) {
        warning('Note that we always assume there are more than one cluster.')
        NUMC <- NUMC[NUMC > 1]
    }

    #Why is this performed here?
    W <- (W + t(W))/2
    diag(W) <- 0

    #NUMC validity check
    if (length(NUMC) <= 0) {
        warning(paste("Invalid NUMC provided, must be an integer vector",
             "with atleast one other number than 1.",
              "Using default NUMC=c(2,3,4,5)",sep=""))
        NUMC <- 2:5 
    }

    # compute unnormalized Laplacian
    degs <- rowSums(W)
    degs[degs == 0] <- .Machine$double.eps    
    D <- diag(degs)    
    L <- D - W
    Di <- diag(1 / sqrt(degs))
    L <- Di %*% L %*% Di
    #print(dim(L))

    # compute the eigenvectors corresponding to the k smallest
    eigs <- eigen(L)
    eigs_order <- sort(eigs$values, index.return=T)$ix
    eigs$values <- eigs$values[eigs_order]
    eigs$vectors <- eigs$vectors[, eigs_order]
    eigengap <- abs(diff(eigs$values))
#    eigengap <- eigengap * (1 - eigs$values[1:length(eigs$values) - 1]
#        ) / (1 - eigs$values[2:length(eigs$values)])

    quality <- list()
    for (c_index in 1:length(NUMC)) {
        ck <- NUMC[c_index]
        UU <- eigs$vectors[, 1:ck]
        EigenvectorsDiscrete <- .discretisation(UU)[[1]]
        EigenVectors <- EigenvectorsDiscrete^2
      
        #MATLAB: sort(EigenVectors,2, 'descend');
        temp1 <- EigenVectors[do.call(order, lapply(1:ncol(EigenVectors),
             function(i) EigenVectors[, i])), ]
        temp1 <- t(apply(temp1, 1, sort, TRUE))  
  
        quality[[c_index]] <- (1 - eigs$values[ck + 1]) / 
            (1 - eigs$values[ck]) * 
            sum( sum( diag(1 / (temp1[, 1] + .Machine$double.eps) ) %*%
            temp1[, 1:max(2, ck-1)] ))
    }
    #Eigen-gap best two clusters
    m1 <- eigengap[NUMC]
    t1 <- sort(eigengap[NUMC], decreasing=TRUE, index.return=T)$ix
    K1 <- NUMC[t1[1]]
    K12 <- NUMC[t1[2]]

    #Rotation cost best two clusters
    m2 <- quality
    t2 <- sort(unlist(quality), index.return=TRUE)$ix
    K2 <- NUMC[t2[1]]
    K22 <- NUMC[t2[2]]    
  
    output <- list("Eigen-gap best"=K1, "Eigen-gap 2nd best"=K12,
        "Rotation cost best"=K2, "Rotation cost 2nd best"=K22)
    output_2 <- list(m1, m2, K1, K12, K2, K22)
    return (output_2)
}
```

## run loop to find best parameters
K <- 10:30
C <- 2:8
A <- c(0.3, 0.4, 0.5, 0.6, 0.7, 0.8)
T <- 10:20


options(warn=-1)
table_measures_2 <- data.frame(cluster = NA, eigen_gap = NA, rotation_cost = NA, K = NA, A = NA, T = NA)

for (k in K) {
  for (a in A) {
    W1 = affinityMatrix(Dist1, k, a)
    W2 = affinityMatrix(Dist2, k, a)
    W3 = affinityMatrix(Dist3, k, a)
    for(t in T){
      W = SNF(list(W1,W2, W3), k, t)
      for (c in C) {
        x <- my_estimateNumberOfClustersGivenGraph(W, NUMC=c:c)
        table_measures <- data.frame(cluster = c, eigen_gap = x[[1]], rotation_cost = unlist(x[[2]]), K = k, A = a, T = t)
        table_measures_2 <- rbind(table_measures, table_measures_2)
        
      }
    }
  }
  print(k)
}

SNF_table <- table_measures_2
SNF_table <- SNF_table[order(SNF_table$eigen_gap, decreasing = TRUE),]
SNF_table$rank_eigen <- 1:nrow(SNF_table)
SNF_table <- SNF_table[order(SNF_table$rotation_cost),]
SNF_table$rank_rot <- 1:nrow(SNF_table)
SNF_table$tot <- SNF_table$rank_rot + SNF_table$rank_eigen
SNF_table_2 <- SNF_table[SNF_table$tot == min(SNF_table$tot),]

## run final model

```{r}
k = 30;##number of neighbors, usually (10~30)
a = 0.7; ##hyperparameter, usually (0.3~0.8)
t = 20; ###Number of Iterations, usually (10~20)
```

```{r}
W1 = affinityMatrix(Dist1, k, a)
W2 = affinityMatrix(Dist2, k, a)
W3 = affinityMatrix(Dist3, k, a)
W = SNF(list(W1,W2, W3), k, t)
```


```{r}
write.csv(W, "processing/SNG_fused_network.csv")
```
```{r}
W <- read.csv("processing/SNG_fused_network.csv")
```

## SNF
```{r}
W = SNF(list(W1,W2, W3), k, t)
c = 2###number of clusters
group = spectralClustering(W,c);###the final subtypes information
```

```{r}
table(group)
```
## extract cluters
```{r}
table_cluters <- data.frame(cluster = group, names = rownames(data_met))
write.csv(table_cluters, "processing/cluster_SNF_transcriptomics_metabolomics_olink_3.csv")
```

```{r}
x <- read.csv("processing/cluster_SNF_transcriptomics_metabolomics_olink_3.csv")
```

```{r}
table_clu <- merge(x, table_cluters, by = "names")
```

```{r}
min(W)
max(W)
```

```{r}
x <- colorRampPalette(c("black", "#007f7f","#00ffff"))(200)
```

```{r}
displayClusters2 <- function(W, group, ColSideColors) {
    # Visualize the clusters present in given similarity matrix with sample 
    #   information displayed by given colors.
    #
    # Args:
    #   W: Affinity matrix
    #   group: labels of cluster groups 
    #   ColSideColors: Character vector of length(group) containing color names 
    #       for horizontal side bar to annotate columns of W, OR a character 
    #       matrix with number of rows matching number of rows in W.
    #
    # Returns:
    #   NULL
    normalize <- function(X) X/rowSums(X)
    ind <- sort(as.vector(group), index.return = TRUE)
    ind <- ind$ix

    diag(W) <- median(as.vector(W))
    W <- normalize(W)
    W <- W + t(W)
    x <- heatmap(W[ind, ind],scale="none",Rowv=NA,Colv=NA, ColSideColors=ColSideColors[ind], labRow = FALSE, labCol = FALSE, col = x)
    return(x)
}
```

```{r}
displayClusters(W1,group)
```

```{r}
displayClusters(W2,group)
```
  
```{r}
color_group <- group
color_group[color_group == 1] <- col_c[1]
color_group[color_group == 2] <-col_c[2]
```

```{r}
pdf("results/figures/Heatmap_transcriptomics_cocomo_SNF_3.pdf", height = 4.5, width = 5)
displayClusters2(W1,group, color_group)
dev.off()
pdf("results/figures/Heatmap_metabolites_cocomo_SNF_3.pdf", height = 4.5, width = 5)
displayClusters2(W2,group, color_group)
dev.off()

pdf("results/figures/Heatmap_olink_cocomo_SNF_3.pdf", height = 4.5, width = 5)
displayClusters2(W3,group, color_group)
dev.off()

pdf("results/figures/Heatmap_whole_cocomo_SNF_3.pdf", height = 4.5, width = 5)
displayClusters2(W,group, color_group)
dev.off()
```

```{r}
displayClusters(W,group);
#SNFNMI = Cal_NMI(group, truelabel)
ConcordanceMatrix = concordanceNetworkNMI(list(W, W1,W2, W3), c);
sum(ConcordanceMatrix)
```

```{r}
write.csv(ConcordanceMatrix, "processing/concordance_matrix_3.csv")
```


```{r}
library(ComplexHeatmap)

pdf("results/figures/NMI_matrix_selection_clusters_3.pdf")
col = colorRamp2(c(0, 0.1, 0.5), c("white", "#9999cc","#000080"))
Heatmap(ConcordanceMatrix, cluster_rows = FALSE, col = col, cluster_columns = FALSE)
dev.off()
```

```{r}
n_clus <- 2:10
x <- my_estimateNumberOfClustersGivenGraph(W, NUMC=n_clus)

table_measures <- data.frame(cluster = n_clus, eigen_gap = x[[1]], rotation_cost = unlist(x[[2]]))
```

## figure eigen gap/ rotation cost
```{r}
scaleFactor <- max(table_measures$eigen_gap) / max(table_measures$rotation_cost)
```

```{r}
ggplot(table_measures, aes(x=cluster)) +
  geom_line(aes(y=eigen_gap), method="loess", col="blue", size=2) +
  geom_line(aes(y=rotation_cost * scaleFactor), method="loess", col="red" , size=2) +
  scale_y_continuous(name="eigen gap", sec.axis=sec_axis(~./scaleFactor, name="rotation cost"))+ 
    theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0)), legend.position = "none") +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  )
ggsave("results/figures/SNF_selection_clusters_3.pdf")
```

```{r}
library(reshape2)
W <- read.csv("/home/flomik/Desktop/Code-PHD/extra/COCOMO_txn_3/processing/SNF_2_cluster_input_Cytoscape_no_filter_3.csv")
W$X <- NULL


W <- dcast(W, Var2 ~ Var1)
W[is.na(W)] <- 1

rownames(W) <- W$Var2
W$Var2 <- NULL

for (i in 1:158){
  for (j in 1:158){
    if(W[i,j] == 0){
      W[i,j] <- W[j,i]
    }
  }
}
```

```{r}
x
```

```{r}
names_a <- x$names[match(colnames(W),x$names)]
group <- x$cluster[match(colnames(W),x$names)]
```

```{r}
colnames(W)
```
```{r}
rownames(W)
```

## PCA
```{r}
pca= prcomp(W) 
df_out <- as.data.frame(pca$x)
df_out$group <- group
```

```{r}
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
barplot(project.pca.proportionvariances, cex.names=1, xlab=paste("Principal component (PC), 1-", length(pca$sdev)), ylab="Proportion of variation (%)", main="Scree plot", ylim=c(0,100))
PCAloadings <- data.frame(Variables = rownames(pca$rotation), pca$rotation)
PCAloadings_2 <- PCAloadings
x <- abs(PCAloadings_2$PC1) < 0.15 & abs(PCAloadings_2$PC2) < 0.15
PCAloadings_2$PC1[x] <- NA
PCAloadings_2$PC2[x] <- NA
PCAloadings_2$Variables[x] <- NA
```
```{r}
col <- c("#878782",	"#761212")
```

```{r}
df_out$group <- factor(df_out$group)
```

```{r}
ggplot(df_out)+
    geom_point(aes(x=PC1,y=PC2, color=group, fill = group), size=4) +
    theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
            legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
            axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
            axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
        theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
              axis.title.y = element_text(size=16, face = "bold",color="black"),
              axis.text = element_text(color="black",size=16, face = "bold"),
              axis.text.x =element_text(color="black",size=16, face = "bold"),
              panel.border = element_blank(),
              axis.ticks.x = element_blank(),
              strip.text.x = element_text(size = 9,color="black"),
              panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
              panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
              panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
              plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
              legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values = col)+ scale_fill_manual(values = col)+ scale_fill_manual(values = col)+ geom_density_2d(aes(x=PC1,y=PC2,color=group, fill = group))+xlim(-0.18, 0.22)+ylim(-0.2, 0.32)

path_fig <- paste0("results/figures/PCA_SNF_fusion_2_density.pdf")
dev.copy(pdf, path_fig, height = 4, width = 6)
dev.off()
```


```{r}
gg <- data.frame(cluster=factor(df_out$group), x=df_out$PC1, y=df_out$PC2)
centroids <- aggregate(cbind(x,y)~cluster,data=gg,mean)
gg <- merge(gg,centroids,by="cluster",suffixes=c("",".centroid"))

ggplot(gg)+
  geom_point(aes(x=x,y=y,color=cluster, fill = cluster), size=3)+
  stat_ellipse(geom = "polygon", aes(x=x,y=y,color=cluster, fill = cluster), alpha = 0.3) +
  geom_point(data=centroids, aes(x=x, y=y, color=cluster, fill = cluster), size = 5, alpha = 0.9, shape = 19)+ 
  geom_segment(aes(x=x.centroid, y=y.centroid, xend=x, yend=y, color=cluster))+
  theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+
    theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0)))+ scale_color_manual(values=col)+ scale_fill_manual(values=col)

path_fig <- paste0("results/figures/PCA_SNF_fusion_2.pdf")
dev.copy(pdf, path_fig, height = 4, width = 6)
dev.off()
```
```{r}
library(Matrix)
colnames(W) <- colnames(Dist1)
rownames(W) <- colnames(Dist1)
W[lower.tri(W)] <- 0 
W <- melt(W)
W <- data.frame(W[W$Var1 != W$Var2,])
write.csv(W, "processing/SNF_2_cluster_input_Cytoscape_no_filter_3.csv")

W <- read.csv("processing/SNF_2_cluster_input_Cytoscape_no_filter_3.csv")
W <- W[W$value > 0.007,]
nrow(W)
write.csv(W, "processing/SNF_2_cluster_input_Cytoscape.csv")
```


