---
title: "NetDx clustering"
output: html_notebook
---

## Aim
Validate clusters using a second tool

## Documentation
https://www.bioconductor.org/packages/release/bioc/vignettes/netDx/inst/doc/ThreeWayClassifier.html#4_Rules_to_create_features_(patient_similarity_networks)

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

### load packages
```{r}
suppressWarnings(suppressMessages(require(netDx)))
library(mygene)
library(org.Hs.eg.db)
library(MultiAssayExperiment)
library(piano)
```

```{r}
col_c <- c("#6b8ea7",	"#a7846b")
```

## load data
### transcriptomics
```{r}
data_txn <- read.csv("processing/COCOMO_transcriptomics_vsd_data_deseq2.csv", row.names = 1, check.names = FALSE)
data_txn <- data.frame(t(data_txn))
data_txn$Gene_name <- mapIds(org.Hs.eg.db, rownames(data_txn), keytype="ENSEMBL", column="SYMBOL")
data_txn <- data_txn[!is.na(data_txn$Gene_name),]
data_txn <- data_txn[!duplicated(data_txn$Gene_name),]
rownames(data_txn) <- data_txn$Gene_name
data_txn$Gene_name <- NULL
data_txn <- data.frame(t(data_txn))
#data_txn$COCOMO_ID <- rownames(data_txn)
rownames(data_txn) <- gsub("X", "", rownames(data_txn))
```
### metabolomics
```{r}
data_met <- read.csv("processing/COCOMO_metabolomics_log2_norm_filt.csv", row.names = 1, check.names = FALSE)
data_met <- data_met[rownames(data_met) %in% rownames(data_txn), ]
data_met <- data_met[order(match(rownames(data_met), rownames(data_txn))),]
#data_met$COCOMO_ID <- rownames(data_met)
```
### olink
```{r}
data_olink <- read.csv("processing/olink_data_new_ids.csv", row.names = 1, check.names = FALSE)
data_olink$Condition <- NULL
```


### clinical parameters
```{r}
clinical <- read.csv("processing/clinical_data_clean_with_clusters_and_categories.csv")
clinical$X <- NULL
clinical$User <- NULL
clinical_HIV <- clinical[!is.na(clinical$METS),]
clinical_HIV <- clinical_HIV[clinical_HIV$COCOMO_ID %in% rownames(data_olink),]
```

```{r}
data_txn <- data_txn[rownames(data_txn) %in% clinical_HIV$COCOMO_ID,]
data_met <- data_met[rownames(data_met) %in% clinical_HIV$COCOMO_ID,]
data_olink <- data_olink[rownames(data_olink) %in% clinical_HIV$COCOMO_ID,]
```

```{r}
data_met <- data_met[order(match(rownames(data_met), rownames(data_txn))),]
data_olink <- data_olink[order(match(rownames(data_olink), rownames(data_txn))),]
```

## filter data
```{r}
data_met <- t(data_met)
data_met <- data_met[apply(data_met, 1, var) >= 0.2, ]

data_olink <- t(data_olink)
data_olink <- data_olink[apply(data_olink, 1, var) >= 0.2, ]

data_txn <- t(data_txn)
data_txn <- data_txn[apply(data_txn, 1, var) >= 0.2, ]
```

```{r}
colnames(data_met) <- paste0("P", colnames(data_met))
colnames(data_txn) <- paste0("P", colnames(data_txn))
colnames(data_olink) <- paste0("P", colnames(data_olink))
```

```{r}
clinical_HIV$cluster <- paste0("C", clinical_HIV$cluster)
```


## check clinical
```{r}
patient.data_2 <- data.frame(
  patientID = paste0("P", clinical_HIV$COCOMO_ID),
  age= clinical_HIV$AGE,
  gender = clinical_HIV$GENDER,
  BMI = clinical_HIV$BMI,
  Duration = clinical_HIV$Duration,
  MetS = clinical_HIV$METS,
  Ethnicity = clinical_HIV$Ethnic,
  SAT = clinical_HIV$SAT,
  VAT = clinical_HIV$VAT,
  CD4_CD8 = clinical_HIV$CDA_CD8,
  ID = paste0("P", clinical_HIV$COCOMO_ID),
  STATUS = clinical_HIV$cluster,
  row.names= paste0("P", clinical_HIV$COCOMO_ID))

Exp <- list("txn"= data_txn, "prot" = data_olink, "met" = data_met)
MultiAssay <- MultiAssayExperiment(experiments=Exp, colData=patient.data_2)
```

## add clinical to groupList
```{r}
expr_2 <- assays(MultiAssay)
groupList_2 <- list()
for (k in 1:length(expr_2)) { # loop over all layers
    cur <- expr_2[[k]]; nm <- names(expr_2)[k]

    # all measures from this layer go into our single PSN
    groupList_2[[nm]] <- list(nm=rownames(cur)) 

    # assign same layer name as in input data
    names(groupList_2[[nm]])[1] <- nm;
}

groupList_2[["clinical"]] <- list(
    age="age",
    MetS = "MetS",
    gender = "gender",
    BMI = "BMI",
    Duration = "Duration",
    Ethnicity = "Ethnicity",
    CD4_CD8 = "CD4_CD8",
    status = "STATUS"
)
```

## load pathways / group genes by pathways
```{r}
gset <- loadGSC("/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/GSEA/h.all.v7.5.1.symbols.gmt")
gset <- gset$gsc
#groupList_2[["txn"]] <- gset
#groupList_2[["prot"]] <- gset

#pathFile <- sprintf("%s/extdata/pathway_ex3.gmt", path.package("netDx"))
#pathList <- suppressMessages(readPathways(pathFile))
#groupList_2[["prot"]] <- pathList
```

## type of comparison
```{r}
sims_2 <- list(a="pearsonCorr", b="pearsonCorr", c = "pearsonCorr", d = "normDiff")
names(sims_2) <- names(groupList_2)
## "normDiff" = clinical data
```

look for features that score 7+ out of 10 in >70% splits
featScoreMax=10, featSelCutoff=7 and numSplits=100L

```{r}
Data <- dataList2List(MultiAssay, groupList_2)

DataList <- Data$assays
DataList <- c(DataList, list(Data$pheno))
names(DataList)[5] <- "pheno"

data <- convertToMAE(DataList)

nco_2 <- round(parallel::detectCores()*0.75) # use 75% available cores
message(sprintf("Using %i of %i cores", nco_2, parallel::detectCores()))
## Using 54 of 72 cores
outDir_2 <- "/home/flomik/Desktop/Code-PHD/COCOMO_txn/results/netDx/netDx_m5"

if (file.exists(outDir_2)) unlink(outDir_2,recursive=TRUE)
numSplits <- 20L

t0_2 <- Sys.time()
set.seed(42) # make results reproducible
model_2 <- suppressMessages(
    buildPredictor(
        dataList=data,          ## your data
        groupList=groupList_2,    ## grouping strategy
        sims = sims_2,
        outDir=outDir_2,          ## output directory
        trainProp=0.8,          ## pct of samples to use to train model in each split
        numSplits=20L,           ## number of train/test splits
        featSelCutoff=7L,       ## threshold for calling something feature-selected
        featScoreMax=10L,        ## max score for feature selection
        numCores=nco_2,           ## set higher for parallelizing
        debugMode=FALSE,
        keepAllData=FALSE,  ## set to TRUE for debugging or low-level files used by the dictor
        logging="none"
  ))
t1_2 <- Sys.time()
print(t1_2-t0_2)
```

```{r}
pdf("results/networks/netDx/results_model_5.pdf")
results <- getResults(model_2,unique(colData(data)$STATUS),
                      featureSelCutoff=2L,featureSelPct=0.50)
dev.off()
```
```{r}
summary(results)
```
```{r}
results$performance
```

```{r}
results$featureScores
```
```{r}
pdf("results/networks/netDx/results_model_5_confusion_matrix.pdf")
confMat <- confusionMatrix(model_2)
dev.off()
```
```{r}
results$selectedFeatures$C1
write.csv(results$selectedFeatures$C1, "results/networks/netDx/results_model_5_features_C1.csv")
write.csv(results$selectedFeatures$C2, "results/networks/netDx/results_model_5_features_C2.csv")
```
```{r}
psn <- suppressMessages(getPSN(
  data,
  groupList_2,
  sims=sims_2,
  selectedFeatures=results$selectedFeatures,
  plotCytoscape = TRUE,
))

write.csv(psn$patientSimNetwork_pruned, "results/networks/netDx/results_model_5_patientSimNetwork_pruned.csv", quote = FALSE)
write.csv(psn$patientSimNetwork_unpruned, "results/networks/netDx/results_model_5_patientSimNetwork_unpruned.csv", quote = FALSE)
```