---
title: "Extract genes database"
output: html_notebook
---


### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/COCOMO_txn/")) 
```

## Genes --> forebrain, human
cerebrum
Thalamus
Hypotalamus

## load packages
```{r loadPackages, include=FALSE, cache=FALSE}
## load additional packages in this chunk
library(pander)
library(knitr)
library(ggplot2)
library(dplyr)
library(igraph)
library(UpSetR)
library(igraph)
```

## database location
```{r}
dbname<-"/home/flomik/Desktop/Code-PHD/COCOMO_txn/data/synaptic.proteome_SR_20210408.db.sqlite"
```

## Connect DB
```{r connect}
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = dbname)
```

# Query
## Get specific genes information
```{r get.specific.genes}
gns <-con %>% tbl("FullGeneFullPaperFullRegion")

gns <- data.frame(gns)
names(gns)

gns <-  select(gns, Localisation, HumanName, BrainRegion, SpeciesTaxID)
```

```{r}
table(gns$SpeciesTaxID)
```
9606 : Human
10090 : Mus musculus Linnaeus
10116 : Rattus norvegicus

## select human genes
```{r}
gns <- gns[gns$SpeciesTaxID == "9606",]
gns$SpeciesTaxID <- NULL
```

```{r}
table(gns$BrainRegion)
```
The outer layer of the cerebrum is called the cerebral cortex, and in each hemisphere it is traditionally divided into four lobes - frontal, parietal, occipital and temporal.

```{r}
write.csv(gns, "data/brain_genes_clean_table.csv")
```

